---
title: "covid 19"
author: "Lelainy Saavedra"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex
$H_0$:There is no correlation between Dying from Covid-19 is and sex.  
$H_A$:There is a correlation between Dying from Covid 19 and sex.  




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  


```{r}
barchartGC(~result + Sex,data=kRdata, type="percent", main ="percent Bar Chart")
```

According to this bar chart, the percentage of females that contracted the virus is higher than males, and the percentage of females dying is less than males.    
The percentage of males that contracted the virus are less than females, but the percentage of males dying of the virus is higher than females.  
  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
On this table it shows that the total population that is alive 56% are female and 44% are male.  
Out of the total population that is dead 36% are female and 64% are male.  

on the column percent out of all the females 98.7% are alive and 1.4% are dead.
Out of all the males 97% are alive and 3% are dead.  


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The odds of dying if you're a male 44/1411  
The odd of dying if you're a female 25/1831  
The odd ratio is (44/1411)(25/1831)    
The odds ratio is 2.3 to 1  

## Conclusions
The P value is 0.0012 which is less than 0.05, therefore we REJECT the null hypothesis, there IS a correlation between dying of Covid-19 and sex.  
